---
title: "Crow Fault Bars"
author: "Justin Mann"
date: "2024-11-12"
output: html_document
---

```{r libraries}
library(tidyverse) #data wrangling and plotting functions
library(glmmTMB) #ZIP & ZINB models
library(DHARMa) #Dispersion and zero inflation test
library(bbmle) #AICtab for model comparison
theme_set(theme_classic())
```


```{r import short form data}
fbs.df <- read.csv("FeatherMaster_1Nov24.csv", h=T)
fbs.df <- fbs.df[,1:15]
fbs.df$ID <- as.factor(fbs.df$ID)
n_occur <- data.frame(table(fbs.df$ID))
n_occur[n_occur$Freq > 1,]#no repeat IDs
fbs.df$LightPlusMedium <- fbs.df$LightBars + fbs.df$MediumBars #a column combining light and medium bars
```


### Sums of light fault bars in urban and rural habitats
```{r Light fb sums by habitat}
sum(fbs.df$LightBars[fbs.df$Habitat=="urban"]) #200 urban light bars
sum(fbs.df$LightBars[fbs.df$Habitat=="rural"]) #300 rural light bars
```


### Gausian linear model for light fault bars. 
This model uses a Gaussian probability distribution which is not appropriate for count data.
```{r light fb gaussian mdl}
summary(lm(data = fbs.df, LightBars ~ Habitat))
```


### Poisson model for light fault bars.
```{r light fb poisson mdl}
light.poisson.mdl <- (glm(data = fbs.df, LightBars ~ Habitat, family = "poisson"))
```


### DHARMa package zero inflation test
Significantly zero inflated
```{r DHARMa zero inflation test}
light.res.sim <- simulateResiduals(light.poisson.mdl)
testZeroInflation(light.res.sim)
```


### DHARMa package overdispersion test
Significantly overdispersed. This means that we should be using a negative binomial distribution which has an overdispersion parameter.
```{r light fb DHARMa overdispersion test}
testOverdispersion(light.res.sim)
```


### Zero-inflated poisson model
```{r light fb ZIP mdl}
light.ZIP.mdl <- glmmTMB(data = fbs.df, LightBars ~ Habitat, ziformula = ~1, family = "poisson")
```

### Light fault bar AIC model comparison
This output suggests that the negative binomial 2 distribution provides the best fit.
```{r light mdl AIC comparison}
light.ZINB1.mdl <- update(light.ZIP.mdl, family = nbinom1)
light.ZINB2.mdl <- update(light.ZIP.mdl, family = nbinom2)
AICtab(light.poisson.mdl, light.ZIP.mdl, light.ZINB1.mdl, light.ZINB2.mdl)
```


### Best fit light fault bar model
```{r light fb ZINB mdl}
summary(light.ZINB2.mdl)
```


### Light fault bar density plot
This plot shows the frequency of light fault bar numbers per feather in the urban and rural samples. The dotted lines are the habitat means.
```{r light fb density plot}
mean(fbs.df$LightBars[fbs.df$Habitat=="urban"])
mean(fbs.df$LightBars[fbs.df$Habitat=="rural"])

Habitat <- c("Urban", "Rural")
Means <- c(2.27, 3.3)
HabitatMeans <- data.frame(Habitat,Means)

LightBarDensity.plot <- ggplot(fbs.df, aes(x=LightBars, fill = Habitat))+
  geom_density(alpha=0.5)+
  geom_vline(data = HabitatMeans, aes(xintercept = Means, color=Habitat),
             linetype="dashed",
             size = 1,
             show.legend = FALSE)+
  scale_fill_manual(values = c("#D95F02", "#1B9E77"))+
  ylab("Frequency")+
  xlab("Number of light fault bars")

LightBarDensity.plot
```


### Sum of medium fault bars in urban and rural habitats
```{r sum of medium fb per habitat}
sum(fbs.df$MediumBars[fbs.df$Habitat=="urban"])
sum(fbs.df$MediumBars[fbs.df$Habitat=="rural"])
```


### Poisson model for medium fault bars
```{r medium fb poisson mdl}
medium.poisson.mdl <- glmmTMB(data = fbs.df, MediumBars ~ Habitat, family = "poisson")
```


### Zero inflation test for medium fault bars
```{r DHARMa medium fb zero inflation test}
medium.res.sim <- simulateResiduals(medium.poisson.mdl)
testZeroInflation(medium.res.sim)
```


### Overdispersion test for medium fault bars
```{r medium fb DHARMa overdispersion test}
testOverdispersion(medium.res.sim)
```


### Medium fault bar ZIP model
```{r medium fb ZIP mdl}
medium.ZIP.mdl <- glmmTMB(data = fbs.df, MediumBars ~ Habitat, ziformula = ~1, family = "poisson")
```


### Medium fault bar AIC model comparison
This output suggests that the negative binomial 1 distribution provides the best fit.
```{r medium mdl AIC comparison}
medium.ZINB1.mdl <- update(medium.ZIP.mdl, family = nbinom1)
medium.ZINB2.mdl <- update(medium.ZIP.mdl, family = nbinom2)
AICtab(medium.poisson.mdl, medium.ZIP.mdl, medium.ZINB1.mdl, medium.ZINB2.mdl)
```


### Best fit medium fault bar model
This model shows significantly fewer medium fault bars in the urban population
```{r medium fb ZINB1 mdl}
summary(medium.ZINB1.mdl)
```


### Sum of strong fault bars in urban and rural habitats
```{r sum of strong fb per habitat}
sum(fbs.df$StrongBars[fbs.df$Habitat=="urban"])
sum(fbs.df$StrongBars[fbs.df$Habitat=="rural"])
```


### Poisson model for strong fault bars
```{r strong fb poisson mdl}
strong.poisson.mdl <- glmmTMB(data = fbs.df, StrongBars ~ Habitat, family = "poisson")
```


### Strong fault bar zero inflation test
Strong fault bar data are not significantly zero-inflated
```{r DHARMa strong fb zero inflation test}
strong.res.sim <- simulateResiduals(strong.poisson.mdl)
testZeroInflation(strong.res.sim)
```


### Overdispersion test for strong fault bars
```{r strong fb DHARMa overdispersion test}
testOverdispersion(strong.res.sim)
```


### Strong fault bar NB1 model
```{r strong fb NB1 mdl}
strong.NB1.mdl <- glmmTMB(data = fbs.df, StrongBars ~ Habitat, family = nbinom1)
```


### Strong bar AIC model comparison
This output suggests that the negative binomial 1 distribution provides the best fit.
```{r strong mdl AIC comparison}
strong.NB2.mdl <- update(strong.NB1.mdl, family = nbinom2)
AICtab(strong.poisson.mdl, strong.NB1.mdl, strong.NB2.mdl)
```


### Best fit strong fault bar model
```{r}
summary(strong.NB1.mdl)
```


### Poisson model for total number of fault bars
```{r total fb poisson mdl}
total.poisson.mdl <- glmmTMB(data = fbs.df, BarsTotal ~ Habitat, family = "poisson")
```


### Total fault bar zero inflation test
```{r DHARMa total fb zero inflation test}
total.res.sim <- simulateResiduals(total.poisson.mdl)
testZeroInflation(total.res.sim)
```


### Overdispersion test for total fault bars
```{r total fb DHARMa overdispersion test}
testOverdispersion(total.res.sim)
```


### Total fault bar ZIP model
```{r total fb ZIP mdl}
total.ZIP.mdl <- glmmTMB(data = fbs.df, BarsTotal ~ Habitat, ziformula = ~1, family = "poisson")
```


### Total fault bar AIC model comparison
```{r total mdl AIC comparison}
total.NB1.mdl <- update(total.ZIP.mdl, family = nbinom1)
total.NB2.mdl <- update(total.ZIP.mdl, family = nbinom2)
AICtab(total.poisson.mdl, total.ZIP.mdl, total.NB1.mdl, total.NB2.mdl)
```


### Best fit total fault bar model
```{r total fb NB2 mdl}
summary(total.NB2.mdl)
```


### Total fault bar density plot
This plot shows the frequency of total fault bar numbers per feather in the urban and rural samples. The dotted lines are the habitat means.
```{r}
mean(fbs.df$BarsTotal[fbs.df$Habitat=="urban"])
mean(fbs.df$BarsTotal[fbs.df$Habitat=="rural"])

TotHabitat <- c("Urban", "Rural")
TotMeans <- c(3.83, 5.42)
TotHabitatMeans <- data.frame(TotHabitat,TotMeans)

TotalBarDensity.plot <- ggplot(fbs.df, aes(x=BarsTotal, fill = Habitat))+
  geom_density(alpha=0.5)+
  geom_vline(data = HabitatMeans, aes(xintercept = TotMeans, color=TotHabitat),
             linetype="dashed",
             size = 1,
             show.legend = FALSE)+
  scale_fill_manual(values = c("#D95F02", "#1B9E77"))+
  ylab("Frequency")+
  xlab("Total number of fault bars")

TotalBarDensity.plot
```


## To do list
### 1. Restrict by feather type (DONE)
###  - Total fault bars
###  - Medium fault bars

### 2. Combine light and medium fault bars

### 3. Feather weight by Habitat for feather type

### 4. Length of fault bar by Habitat


```{r lpm poisson mdl}
lpm.poisson.mdl <- glmmTMB(data = fbs.df, LightPlusMedium ~ Habitat, family = "poisson")
```


### LightPlusMedium fault bar zero inflation test
```{r DHARMa lpm fb zero inflation test}
lpm.res.sim <- simulateResiduals(lpm.poisson.mdl)
testZeroInflation(lpm.res.sim)
```


### Overdispersion test for lpm fault bars
```{r lpm fb DHARMa overdispersion test}
testOverdispersion(lpm.res.sim)
```


### LightPlusMedium ZIP model
```{r lpm ZIP mdl}
lpm.ZIP.mdl <- glmmTMB(data = fbs.df, LightPlusMedium ~ Habitat, ziformula = ~1, family = "poisson")
```


### LPM fault bar AIC model comparison
```{r lpm mdl AIC comparison}
lpm.NB1.mdl <- update(lpm.ZIP.mdl, family = nbinom1)
lpm.NB2.mdl <- update(lpm.ZIP.mdl, family = nbinom2)
AICtab(lpm.poisson.mdl, lpm.ZIP.mdl, lpm.NB1.mdl, lpm.NB2.mdl)
```


### Best fit LPM model
```{r lpm NB1 mdl}
summary(lpm.NB1.mdl)
```



## Subsetting by feather type

### Primary feather subset
```{r primary df subset}
fbs.prim.df <- subset(fbs.df, Type == "primary")
```


```{r primary total fb ZIP mdl}
summary(glmmTMB(data = fbs.prim.df, BarsTotal ~ Habitat, ziformula = ~1, family = "poisson"))
```


### Primary feather light fault bar Poisson model
Significantly fewer in urban habitat
```{r primary light fb poisson mdl}
prim.light.poisson.mdl <- glmmTMB(data = fbs.prim.df, LightBars ~ Habitat, family = "poisson")
summary(prim.light.poisson.mdl)
```

### Zero inflation test for primary feather light fault bars
Still strongly zero-inflated data
```{r primary light zero inflation test}
#zero inflation test from: http://data.princeton.edu/wws509/r/overdispersion.html 
zobs <- fbs.prim.df$LightBars == 0
zpoi <- exp(-exp(predict(prim.light.poisson.mdl)))
c(obs=mean(zobs), poi = mean(zpoi))
```

### Primary feather light fault bar ZIP model
Nonsignificant relationship
```{r primary light fb ZIP mdl}
prim.light.ZIP.mdl <- glmmTMB(data = fbs.prim.df, LightBars ~ Habitat, ziformula = ~1, family = "poisson")
summary(prim.light.ZIP.mdl)
```

### Primary feather medium fault bar ZIP model
Nonsignificant relationship
```{r primary medium fb mdl}
summary(glmmTMB(data = fbs.prim.df, MediumBars ~ Habitat, ziformula = ~1, family = "poisson"))
```


### Secondary feather subset
```{r secondary df subset}
fbs.sec.df <- subset(fbs.df, Type == "secondary")
```


### Secondary feather light fault bar ZIP model
Nonsignificant relationship
```{r secondary light fb mdl}
summary(glmmTMB(data = fbs.sec.df, LightBars ~ Habitat, ziformula = ~1, family = "poisson"))
```


### Secondary feather medium fault bar ZIP model
Significantly MORE medium fault bars in the secondary feathers
```{r secondary medium fb mdl}
summary(glmmTMB(data = fbs.sec.df, MediumBars ~ Habitat, ziformula = ~1, family = "poisson"))
```

### Tail feather subset
```{r tail df subset}
fbs.tail.df <- subset(fbs.df, Type == "tail")
```


### Tail feather light fault bar ZIP model
Significantly fewer light fault bars in the urban habitat
```{r tail light fb ZIP mdl}
summary(glmmTMB(data = fbs.tail.df, LightBars ~ Habitat, ziformula = ~1, family = "poisson"))
```

### Tail feather medium fault bar ZIP model
Significantly fewer medium fault bars in urban habitat
```{r tail medium fb mdl}
summary(glmmTMB(data = fbs.tail.df, MediumBars ~ Habitat, ziformula = ~1, family = "poisson"))
```


