---
title: "Crow Fault Bars v2"
author: "Justin Mann"
date: "`r Sys.Date()`"
output: html_document
---

```{r libraries, echo=FALSE}
library(tidyverse) #data wrangling and plotting functions
library(glmmTMB) #ZIP & ZINB models
library(DHARMa) #Dispersion and zero inflation test
library(bbmle) #AICtab for model comparison
library(multcomp) #glht() for multiple comparisons
theme_set(theme_classic())
```


```{r}
palette.colors(palette = "Tableau 10")
```



```{r import data}
fb.df <- read.csv("FaultBarMaster_15Nov24.csv", h=T)
fb.df$ID <- as.factor(fb.df$ID)
fb.df$Habitat <- as.factor(fb.df$Habitat)
n_occur <- data.frame(table(fb.df$ID))
n_occur[n_occur$Freq > 1,]#no repeat IDs
```


### Total fault bars by feather type model
```{r total fb by ftype mdl}
tot.type.mdl <- glmmTMB(data = fb.df, TotalBars ~ Type, family = "poisson")
summary(tot.type.mdl)
```


### Zero inflation test for tot.type.mdl
```{r DHARMa tot.type.mdl zero inflation test}
tot.type.res.sim <- simulateResiduals(tot.type.mdl)
testZeroInflation(tot.type.res.sim)
```


### Overdispersion test for tot.type.mdl
```{r tot.type.mdl DHARMa overdispersion test}
testOverdispersion(tot.type.res.sim)
```


### ZIP tot.type.mdl
```{r ZIP tot.type.mdl}
tot.type.ZIP.mdl <- glmmTMB(data = fb.df, TotalBars ~ Type, ziformula = ~1, family = "poisson")
```


### tot.type.mdl AIC comparison
```{r tot.type.mdl AIC comparison}
tot.type.NB1.mdl <- update(tot.type.ZIP.mdl, family = nbinom1)
tot.type.NB2.mdl <- update(tot.type.ZIP.mdl, family = nbinom2)
AICtab(tot.type.mdl,tot.type.ZIP.mdl,tot.type.NB1.mdl,tot.type.NB2.mdl)
```


### tot.type best fit model
```{r tot.type best fit mdl}
summary(tot.type.NB2.mdl)
```



### Number of fault bars (>0) per feather type
```{r TotalBars by Type boxplot}
fb.nozero.df <- fb.df %>% filter(TotalBars>0)
ggplot(data = fb.nozero.df, aes(x=Type, y=TotalBars))+
  geom_boxplot(fill = palette.colors(n = 3, palette = "Tableau 10"))+
  ylab("Total number of fault bars")
```

### Number of light fault bars (>0) per feather type
```{r LightBars by Type boxplot}
fb.light.nozero.df <- fb.df %>% filter(LightBars>0)
ggplot(data = fb.light.nozero.df, aes(x=Type, y=LightBars))+
  geom_boxplot(fill = palette.colors(n = 3, palette = "Tableau 10"))+
  ylab("Number of light fault bars")
```


### Number of strong fault bars (>0) per feather type
```{r StrongBars by Type boxplot}
fb.strong.nozero.df <- fb.df %>% filter(StrongBars>0)
ggplot(data = fb.strong.nozero.df, aes(x=Type, y=StrongBars))+
  geom_boxplot(fill = palette.colors(n = 3, palette = "Tableau 10"))+
  ylab("Number of strong fault bars")
```



## MODELING TOTAL FAULT BARS

### Poisson model for total number of fault bars
```{r total fb poisson mdl}
total.poisson.mdl <- glmmTMB(data = fb.df, TotalBars ~ Habitat, family = "poisson")
```


### Zero inflation test for total.poisson.mdl
```{r DHARMa total fb zero inflation test}
total.res.sim <- simulateResiduals(total.poisson.mdl)
testZeroInflation(total.res.sim)
```


### Overdispersion test for total.poisson.mdl
```{r total fb DHARMa overdispersion test}
testOverdispersion(total.res.sim)
```


### Total fault bar zero inflated poisson model
```{r total fb ZIP mdl}
total.ZIP.mdl <- glmmTMB(data = fb.df, TotalBars ~ Habitat, ziformula = ~1, family = "poisson")
```


### Total fault bar model AIC comparison
```{r total mdl AIC comparison}
total.NB1.mdl <- update(total.ZIP.mdl, family = nbinom1)
total.NB2.mdl <- update(total.ZIP.mdl, family = nbinom2)
AICtab(total.poisson.mdl, total.ZIP.mdl, total.NB1.mdl, total.NB2.mdl)
```


### Best fit total fault bar model
As per AIC comparison, the negative binomial 2 model provides the best fit.
```{r total fb NB2 mdl}
summary(total.NB2.mdl)
```

### Outlier test for total.nb2.mdl
```{r total fb NB2.mdl outlier test}
total.nb2.res.sim <- simulateResiduals(total.NB2.mdl)
testOutliers(total.nb2.res.sim, alternative = "greater")
```


### Total fault bar density plot
This plot shows the frequency of total fault bar numbers per feather in the urban and rural populations. Dotted lines are the urban and rural means.
```{r total fb density plot}
TotHabitat <- c("Urban", "Rural")
TotMeans <- c(mean(fb.df$TotalBars[fb.df$Habitat=="urban"]), mean(fb.df$TotalBars[fb.df$Habitat=="rural"]))
TotHabitatMeans <- data.frame(TotHabitat,TotMeans)

TotalBarDensity.plot <- ggplot(fb.df, aes(x=TotalBars, fill = Habitat))+
  geom_density(alpha = 0.5)+
  geom_vline(data = TotHabitatMeans, aes(xintercept = TotMeans, color = TotHabitat),
             linetype="dashed",
             size = 1,
             show.legend = FALSE)+
  scale_fill_manual(values = c("#F28E2B", "#4E79A7"))+
  ylab("Frequency")+
  xlab("Total number of fault bars (Light + Strong)")

TotalBarDensity.plot
```


## MODELING LIGHT FAULT BARS

### Light fault bar poisson model
```{r light poisson mdl}
light.poisson.mdl <- glmmTMB(data = fb.df, LightBars ~ Habitat, family = "poisson")
summary(light.poisson.mdl)
```


### Zero inflation test for light.poisson.mdl
```{r DHARMa light fb zero inflation test}
light.res.sim <- simulateResiduals(light.poisson.mdl)
testZeroInflation(light.res.sim)
```


### Overdispersion test for light.poisson.mdl
```{r light fb DHARMa overdispersion test}
testOverdispersion(light.res.sim)
```


### Light fault bar zero inflated poisson model
```{r light ZIP mdl}
light.ZIP.mdl <- glmmTMB(data = fb.df, LightBars ~ Habitat, ziformula = ~1, family = "poisson")
```


### Light fault bar model AIC comparison
```{r light mdl AIC comparison}
light.NB1.mdl <- update(light.ZIP.mdl, family = nbinom1)
light.NB2.mdl <- update(light.ZIP.mdl, family = nbinom2)
AICtab(light.poisson.mdl, light.ZIP.mdl, light.NB1.mdl, light.NB2.mdl)
```


### Best fit light model
As per AIC comparison, negative binomial 1 model provides the best fit
```{r light NB1 mdl}
summary(light.NB1.mdl)
```


### Overdispersion test for light.NB1.mdl
```{r light fb DHARMa overdispersion test NB1 mdl}
light.NB1.res.sim <- simulateResiduals(light.NB1.mdl)
testOverdispersion(light.NB1.res.sim)
```


### Outlier test for light.NB1.mdl
The NB1 model has no outliers
```{r light fb DHARMa outlier test}
testOutliers(light.NB1.mdl, alternative = "greater")
```



### Light fault bar density plot
This plot shows the frequency of light fault bar numbers in the urban and rural populations. Dotted lines are the urban and rural means.
```{r light fb density plot}
LightHabitat <- c("Urban", "Rural")
LightMeans <- c(mean(fb.df$LightBars[fb.df$Habitat=="urban"]),mean(fb.df$LightBars[fb.df$Habitat=="rural"]))
LightHabitatMeans <- data.frame(LightHabitat,LightMeans)

LightBarDensity.plot <- ggplot(fb.df, aes(x=LightBars, fill = Habitat))+
  geom_density(alpha=0.5)+
  geom_vline(data = LightHabitatMeans, aes(xintercept = LightMeans, color = LightHabitat),
             linetype="dashed",
             size = 1,
             show.legend = FALSE)+
  scale_fill_manual(values = c("#F28E2B", "#4E79A7"))+
  ylab("Frequency")+
  xlab("Number of light fault bars")

LightBarDensity.plot
```


## MODELING STRONG FAULT BARS

### Strong fault bar poisson model
```{r strong poisson mdl}
strong.poisson.mdl <- glmmTMB(data = fb.df, StrongBars ~ Habitat, family = "poisson")
summary(strong.poisson.mdl)
```


### Zero inflation test for strong.poisson.mdl
```{r DHARMa strong fb zero inflation test}
strong.res.sim <- simulateResiduals(strong.poisson.mdl)
testZeroInflation(strong.res.sim)
```


### Overdispersion test for strong.poisson.mdl
```{r strong fb DHARMa overdispersion test}
testOverdispersion(strong.res.sim)
```


### Strong fault bar zero inflated poisson model
```{r strong ZIP mdl}
strong.ZIP.mdl <- glmmTMB(data = fb.df, StrongBars ~ Habitat, ziformula = ~1, family = "poisson")
```

```{r strong NB1 mdl}
strong.NB1.mdl <- glmmTMB(data = fb.df, StrongBars ~ Habitat, ziformula = ~1, family = "nbinom1")
```

```{r strong NB2 mdl}
strong.NB2.mdl <- glmmTMB(data = fb.df, StrongBars ~ Habitat, ziformula = ~1, family = "nbinom2")
```



### Strong fault bar model AIC comparison
```{r strong mdl AIC comparison}
AICtab(strong.poisson.mdl, strong.ZIP.mdl, strong.NB1.mdl)
```


### Best fit strong model
AIC comparison says that ZIP model fits best, but the zero inflation test is non significant. There is not a strong difference between the ZIP and NB1 model, so I'm including the NB1 model result. The result is non significant but shows a trend that's opposite of the light model; there are marginally MORE strong bars in urban habitats.
```{r best fit strong mdl}
summary(strong.NB1.mdl)
```


### Overdispersion test for strong.NB1.mdl
```{r strong fb DHARMa overdispersion test NB1 mdl}
strong.ZIP.res.sim <- simulateResiduals(strong.ZIP.mdl)
testOverdispersion(strong.ZIP.res.sim)
```


### Outlier test for strong.NB1.mdl
The NB1 model has no outliers
```{r strong fb DHARMa outlier test}
testOutliers(strong.NB1.mdl, alternative = "greater")
```



### Strong fault bar density plot
This plot shows the frequency of strong fault bar numbers in the urban and rural populations. Dotted lines are the urban and rural means.
```{r strong fb density plot}
StrongHabitat <- c("Urban", "Rural")
StrongMeans <- c(mean(fb.df$StrongBars[fb.df$Habitat=="urban"]),mean(fb.df$StrongBars[fb.df$Habitat=="rural"]))
StrongHabitatMeans <- data.frame(StrongHabitat,StrongMeans)

StrongBarDensity.plot <- ggplot(fb.df, aes(x=StrongBars, fill = Habitat))+
  geom_density(alpha=0.5)+
  geom_vline(data = StrongHabitatMeans, aes(xintercept = StrongMeans, color = StrongHabitat),
             linetype="dashed",
             size = 1,
             show.legend = FALSE)+
  scale_fill_manual(values = c("#F28E2B", "#4E79A7"))+
  ylab("Frequency")+
  xlab("Number of strong fault bars")

StrongBarDensity.plot
```



## MODELING TAIL FAULT BARS

### Tail feather subset
```{r tail df subset}
fb.tail.df <- subset(fb.df, Type == "tail")
```


### Total tail fault bar poisson model
```{r tot tail poisson mdl}
tot.tail.poisson.mdl <- glmmTMB(data = fb.tail.df, TotalBars ~ Habitat, family = "poisson")
```


### Zero inflation test for tot.tail.poisson.mdl
```{r DHARMa tot.tail fb zero inflation test}
tot.tail.res.sim <- simulateResiduals(tot.tail.poisson.mdl)
testZeroInflation(tot.tail.res.sim)
```


### Overdispersion test for tot.tail.poisson.mdl
```{r tot.tail fb DHARMa overdispersion test}
testOverdispersion(tot.tail.res.sim)
```


### Total tail fault bar zero inflated poisson model
```{r tot.tail ZIP mdl}
tot.tail.ZIP.mdl <- glmmTMB(data = fb.tail.df, TotalBars ~ Habitat, ziformula = ~1, family = "poisson")
```


### Total tail fault bar model AIC comparison
```{r tot.tail mdl AIC comparison}
tot.tail.NB1.mdl <- update(tot.tail.ZIP.mdl, family = nbinom1)
tot.tail.NB2.mdl <- update(tot.tail.ZIP.mdl, family = nbinom2)
AICtab(tot.tail.ZIP.mdl, tot.tail.NB1.mdl, tot.tail.NB2.mdl)
```


### Best fit tot.tail model
As per AIC comparison, negative binomial 2 model provides the best fit
```{r tot.tail NB2 mdl}
summary(tot.tail.NB2.mdl)
```


### Light tail fault bar poisson model
```{r light tail poisson mdl}
light.tail.poisson.mdl <- glmmTMB(data = fb.tail.df, LightBars ~ Habitat, family = "poisson")
```


### Zero inflation test for light.tail.poisson.mdl
```{r DHARMa light.tail fb zero inflation test}
light.tail.res.sim <- simulateResiduals(light.tail.poisson.mdl)
testZeroInflation(light.tail.res.sim)
```


### Overdispersion test for light.tail.poisson.mdl
```{r light.tail fb DHARMa overdispersion test}
testOverdispersion(light.tail.res.sim)
```


### Light tail fault bar zero inflated poisson model
```{r light.tail ZIP mdl}
light.tail.ZIP.mdl <- glmmTMB(data = fb.tail.df, LightBars ~ Habitat, ziformula = ~1, family = "poisson")
```


### Light tail fault bar model AIC comparison
```{r light.tail mdl AIC comparison}
light.tail.NB1.mdl <- update(light.tail.ZIP.mdl, family = nbinom1)
light.tail.NB2.mdl <- update(light.tail.ZIP.mdl, family = nbinom2)
AICtab(light.tail.ZIP.mdl, light.tail.NB1.mdl, light.tail.NB2.mdl)
```


### Best fit light.tail model
As per AIC comparison, negative binomial 2 model provides the best fit
```{r light.tail NB2 mdl}
summary(light.tail.NB2.mdl)
```



## MODELING SECONDARY FAULT BARS

### Secondary feather subset
```{r secondary df subset}
fb.sec.df <- subset(fb.df, Type == "secondary")
```


### Total secondary fault bar poisson model
```{r tot sec poisson mdl}
tot.sec.poisson.mdl <- glmmTMB(data = fb.sec.df, TotalBars ~ Habitat, family = "poisson")
```


### Zero inflation test for tot.sec.poisson.mdl
```{r DHARMa tot.sec fb zero inflation test}
tot.sec.res.sim <- simulateResiduals(tot.sec.poisson.mdl)
testZeroInflation(tot.sec.res.sim)
```


### Overdispersion test for tot.sec.poisson.mdl
```{r tot.sec fb DHARMa overdispersion test}
testOverdispersion(tot.sec.res.sim)
```


### Total secondary fault bar zero inflated poisson model
```{r tot.sec ZIP mdl}
tot.sec.ZIP.mdl <- glmmTMB(data = fb.sec.df, LightBars ~ Habitat, ziformula = ~1, family = "poisson")
```


### Total tail fault bar model AIC comparison
```{r tot.sec mdl AIC comparison}
tot.sec.NB1.mdl <- update(tot.sec.ZIP.mdl, family = nbinom1)
tot.sec.NB2.mdl <- update(tot.sec.ZIP.mdl, family = nbinom2)
AICtab(tot.sec.poisson.mdl, tot.sec.ZIP.mdl, tot.sec.NB1.mdl, tot.sec.NB2.mdl)
```


### Best fit tot.sec model
As per AIC comparison, negative binomial 1 model provides the best fit.
Result is non significant
```{r tot.sec NB1 mdl}
summary(tot.sec.NB1.mdl)
```



## MODELING PRIMARY FAULT BARS

### Primary feather subset
```{r prim df subset}
fb.prim.df <- subset(fb.df, Type == "primary")
```


### Total primary fault bar poisson model
```{r tot prim poisson mdl}
tot.prim.poisson.mdl <- glmmTMB(data = fb.prim.df, TotalBars ~ Habitat, family = "poisson")
```


### Zero inflation test for tot.prim.poisson.mdl
```{r DHARMa tot.prim fb zero inflation test}
tot.prim.res.sim <- simulateResiduals(tot.prim.poisson.mdl)
testZeroInflation(tot.prim.res.sim)
```


### Overdispersion test for tot.prim.poisson.mdl
```{r tot.prim fb DHARMa overdispersion test}
testOverdispersion(tot.prim.res.sim)
```


### Total prim fault bar zero inflated poisson model
```{r tot.prim ZIP mdl}
tot.prim.ZIP.mdl <- glmmTMB(data = fb.prim.df, LightBars ~ Habitat, ziformula = ~1, family = "poisson")
```


### Total prim fault bar model AIC comparison
```{r tot.prim mdl AIC comparison}
tot.prim.NB1.mdl <- update(tot.prim.ZIP.mdl, family = nbinom1)
tot.prim.NB2.mdl <- update(tot.prim.ZIP.mdl, family = nbinom2)
AICtab(tot.prim.poisson.mdl, tot.prim.ZIP.mdl, tot.prim.NB1.mdl, tot.prim.NB2.mdl)
```


### Best fit tot.prim model
As per AIC comparison, negative binomial 1 model provides the best fit
```{r tot.prim NB1 mdl}
summary(tot.prim.NB1.mdl)
```
